Subset coercions in \Coq

### 2006-07-21 papers/article/main.tex
Continue article

@@

\subsection{Explicit coercions}
The derivation of explicit coercions (figure
\vref{fig:coerce-impl-rules-short}) is based on the
algorithmic equivalence derivations, hence we can derive a function from
the judgement.

@@

\subsection{Terms}
- Follows algo typing derivation. Done at the same time as typing in the
impl.

@@

>>>>

\def\sidecond{\sideconden}

@@

\subsection{Explicit coercions}
The derivation of explicit coercions (figure
\vref{fig:coerce-impl-rules-short}) is based on the
algorithmic equivalence derivations, hence we can derive a function from
the judgement which build a coercion given two types. 
The side-conditions of the rules have not changed, so we omit them for
improved readability.

@@

Coercions are formalized as multi-holes evaluation contexts, with
$\ctxdot$ the denotation of a hole. We define the instanciation $C[x]$ of a context $C$
by a term $x$ as simultaneous substitution of $x$ for every $\ctxdot$ in $C$.
Hence if $C$ is an evaluation context and $x$ a term, $C[x]$ is a term
(ie, it cannot contain $\ctxdot$).
The holes will denote the object to which the coercion is applied. Other
presentations are possible, notably the ``\LMS{}'' style of \cite{longo95logic} which
was adopted by Gang Chen in his treatment of coercive subtyping for the
Calculus of Constructions \cite{conf/popl/Chen03}, and the presentation 
of Amokrane Sa\"ibi \cite{saibi97inheritance}. The \LMS{} style consists
in defining a judgement $x : A \type y : B$ which reads: $A$ is
coercible to $B$ and the coercion is $\lambda x : A.y$. This
presentation is very elegant and well suited for checking coherence and
of the coercion judgement, which we don't need here as we have unicity
of coercions. Moreover, it did not seem readily usable in
our setting, where coercions may contain metavariables. Similarly, Amokrane Sa\"ibi defines
a coercion derivation algorithm which respect some coherence criteria,
roughly, that type-erased coercions $\beta\eta$-reduce to the
identity. Again this wasn't adapted to our setting, so we invented this
new presentation.

We suppose that conversion is preserved by the interpretation $\ip{}{}$, so
the rule \irule{SubConv} derives an empty coercion, as the target system
will be able to derive itself that $\ip{T}{`G} `= \ip{U}{`G}$.
Similarly for \irule{SubHnf}, we rely on the target's system conversion
rule so that the coercion $c$ of codomain $\ip{\hnf{U}}{`G}$ can be seen
as an object of type $\ip{U}{`G}$.

Next, we have the rules for products and sums, which compose
coercions. In the \irule{SubProd} rule, we first coerce the argument to
an object of type $\ip{T}{`G}$. Then we apply it to the coerced
function, resulting in an object $\ctxdot~(c_1[x])$ of type
$\ip{V}{`G, x : T}[c_1[x]/x]$. We'll see later that this type is equivalent to
$\ip{V}{`G, x : U}$, so we can apply the second coercion and get an
object of type $\ip{W}{`G}$. Similarly for dependent sums \TODO{}

Now for the original part, the \irule{SubSub} rule simply inserts a
projection before applying the inductively defined coercion on the
support of the subset. On the other hand, \irule{SubProof} creates and
object of a subset using its sole constructor \sref{elt}. The object
part is given by the coercion between the support while the proof part
is promised using a metavariable of type $\ip{P}{`G, x : U}[c/x]$ which
we will turn later into a proof obligation.

\subsection{Interpretation of terms}
The interpretation of terms from \Russell to \CCq (figure
\vref{fig:interp}) is a straightforward
recursive traversal of algorithmic typing derivations. In other, more
practical words, we can do interpretation of terms simultaneously with
typing in our implementation, just like for the existing coercion system
of \Coq.

\interpfig[Interpretion of terms]

We describe the case of application here, others are similar. First we
get the types $F$ and $U$ of the function $f$ and argument $u$ by recursively calling the
typing function on both terms. Then we must ensure that $F$ can indeed
be seen as a product using the $\mualgo$ operator. We can then build
the coercion between $F$ and this product, and between $U$ and its
domain. We finally instantiate these coercions by their corresponding
interpreted objects and return their application.

\subsection{A little more expressiveness}
One may wonder how can we interpret \Russell terms to a system with only
$\beta\pi$-equivalence. Indeed, if the algorithmic coercion judgement is transitive
and has a unicity property, then we may suppose that the coercions
themselves have corresponding properties:
\begin{proposition}[Coercion transitivity]
  If $\subimpl{`G}{c_1}{T}{U}{s}$ and $\subimpl{`G}{c_2}{U}{V}{s}$ then
  there exists a coercion $c$ so that $\subimpl{`G}{c}{T}{V}{s}$ and $c
  \eqqbp c_2 `o c_1$ (where $c_2 `o c_1 = c_2[c_1]$).
\end{proposition}

Now take the following example
\begin{example}[Transitivity needs eta]
  Let $T, U, V = \Pi x : A.B$.
  Then $c_1$ (and $c_2$ by unicity) must have the form $\lambda x :
  \ip{C}{`G}.c^2_1[\ctxdot~c^1_1[x]]$ (respectively $\lambda x : \ip{A}{`G}.c^2_2[\ctxdot~c^1_2[x]]$).
  So $c_2 `o c_1 = \lambda x : \ip{A}{`G}. c^2_1[(\lambda x :
  \ip{A}{`G}.c^2_2[\ctxdot~c^1_2[x]])~c^1_1[]] \redbp \lambda x :
  \ip{A}{`G}. c^2_1[c^2_2[\ctxdot~c^1_2[c^1_1[x]]]]$.
  Suppose all $c^i_j$ have been derived by \irule{SubConv}. Then we have 
  $c_2 `o c_1 = $.

\end{example}

\ccieqfig[Equational theory of \CCq]

@@

\newpage
